MCTS Based on Simple Regret

### 2012-01-21 paper/rmcts/aaai/paper.tex
in progress
added formulas for VOI estimates

\begin{dfn}
The \textbf{simple regret} of a sampling policy for the Multi-armed Bandit
Problem is the expected difference between the best expected reward
$\mu_*$ and the expected reward $\mu_j$ of the empirically best arm
$\overline X_j=\max_i\overline X_i$:
\begin{equation}
\label{eq:simple-regret}
\IE[R]=\sum_{j=1}^K\Delta_j\Pr(\overline X_j=\max_i\overline X_i)
\end{equation}
where $\Delta_j=\mu_*-\mu_j$.
\end{dfn}

@@

The UCT sampling scheme \cite{Kocsis.uct} is a widely used generalization
of UCB for MCTS.  UCT applies the UCB sampling
scheme at each step of a rollout.  We suggest an improvement on UCT,
which combines different sampling schemes on the first step and during
the rest of a rollout:
\begin{dfn}
The {\bf two-stage MCTS sampling scheme} selects an action at the
current root node according to an algorithm that minimizes the simple
regret, such as $\frac 1 2$-greedy or $UCB_{\sqrt{\cdot}}$, and then
selects actions according to $UCB$.
\end{dfn}

@@

A further improvement can be achieved by computing or estimating the VOI
of the rollouts and choosing a rollout that maximizes the VOI. VOI of
a rollout can be computed when the sample distribution of an
action is known up to the parameters, such as the normal distribution
with an unknown mean and/or variance. Alternatively, the value of
information can be estimated from the set of samples, and the need to
assume a particular shape of the distribution can be lifted. In one
realization of the latter approach the VOI of performing an action is
estimated as an \emph{upper bound on the value of perfect information about
an arm, divided by the number of pulls of the arm}. 
\begin{equation}
\end{equation}
Early experiments with this approach demonstrated significantly lower
simple regret on a wide range of problems.

@@

\begin{figure}[t]
  \begin{minipage}[c]{1.0\linewidth}
    \centering
    \includegraphics[scale=1.0]{onelevel-tree.pdf}\\
    a. search tree
    \vspace{1em}
  \end{minipage}
  \begin{minipage}[c]{1.0\linewidth}
    \centering
    \includegraphics[scale=0.5]{flat-trilevel-k=64-uqb=8.pdf}\\
    b. regret vs. number of samples
  \end{minipage}
  \caption{Simple regret in MAB}
  \label{fig:mab-simple-regret}
\end{figure}

@@

\begin{figure}[t]
  \begin{minipage}[b]{1.0\linewidth}
    \centering
    \includegraphics[scale=0.45]{costs-size=6-group=median.pdf}\\
    a. median cost
    \vspace{1em}
  \end{minipage}
  \begin{minipage}[b]{1.0\linewidth}
    \centering
    \includegraphics[scale=0.45]{costs-size=6-group=minimum.pdf}\\
    b. minimum cost
  \end{minipage}
  \caption{The sailing domain, $6\times 6$ lake, cost vs. number of rollouts}
  \label{fig:sailing-cost-vs-nsamples}
\end{figure}

@@

\begin{figure}[t]
  \centering
  \includegraphics[scale=0.45]{costs-size=6-nsamples=397.pdf}\\
  a. 397 rollouts\\
  \vspace{1em}
  \includegraphics[scale=0.45]{costs-size=6-nsamples=1585.pdf}\\
  c. 1585 rollouts
  \caption{The sailing domain, $6\times 6$ lake, cost vs. factor}
  \label{fig:sailing-cost-vs-factor}
\end{figure}

Figure~\ref{fig:sailing-lake-size} shows the cost vs. the exploration
factor for lakes of different sizes. The relative difference between
the sampling schemes becomes more prominent when the lake size
increases.
\begin{figure}[t]
   \centering
   \includegraphics[scale=0.45]{costs-size=6-nsamples=397.pdf}\\
   b. $6\times 6$ lake \\
   \vspace{1em}
   \includegraphics[scale=0.45]{costs-size=10-nsamples=397.pdf}\\
   c. $10\times 10$ lake
  \caption{The sailing domain, 397 samples, cost vs. factor}
  \label{fig:sailing-lake-size}
\end{figure}

Figure~\ref{fig:sailing-rcq-vs-factor} compares QCT with UCT with a
different exploration factor at the root (CCT). $\alpha$ for the rest of
the steps was chosen to ensure the best performance from earlier
experiments on the domain. Both algorithms exhibit similar dependency,
but as the number of samples grows, QCT achieves smaller average
regrets and is less sensitive to the choice of the value for $\alpha$ at
the first step.
\begin{figure}[t]
  \centering
  \includegraphics[scale=0.45]{rcq-size=6-nsamples=397.pdf}\\
  b. 397 rollouts\\
  \vspace{1em}
  \includegraphics[scale=0.45]{rcq-size=6-nsamples=1585.pdf}\\
  d. 1585 rollouts\\
  \caption{The sailing domain, log vs. sqrt, $6\times 6$ lake}
  \label{fig:sailing-rcq-vs-factor}
\end{figure}

@@

The Monte Carlo tree search algorithms presented in the paper differ
from UCT at the first step of the rollout, when the `simple' selection
regret is minimized instead of the cumulative regret. Both the
theoretical analysis and the empirical evaluation provide evidence for
better general performance of the proposed algorithms.

>>>>

\begin{dfn}
The \textbf{simple regret} of a sampling policy for the Multi-armed Bandit
Problem is the expected difference between the best expected reward
$\mu_*$ and the expected reward $\mu_j$ of the empirically best arm
$\overline X_j=\max_i\overline X_i$:
\begin{equation}
\label{eq:simple-regret}
\IE r=\sum_{j=1}^K\Delta_j\Pr(\overline X_j=\max_i\overline X_i)
\end{equation}
where $\Delta_j=\mu_*-\mu_j$.
\end{dfn}

@@

UCT \cite{Kocsis.uct} is a generalization of UCB for MCTS.  UCT
applies UCB at each step of a rollout.  We suggest an improvement on
UCT, which combines different sampling schemes on the first step and
during the rest of a rollout:
\begin{dfn}
The \textbf{SR+CR MCTS sampling scheme} selects an action at the
current root node according to a scheme suitable for minimizing 
simple regret (\textbf{SR}), such as $\frac 1 2$-greedy or $UCB_{\sqrt{\cdot}}$, and
then selects actions according to $UCB$, minimizing cumulative regret (\textbf{CR}).
\end{dfn}

@@

A further improvement can be achieved by computing or estimating the
value of information (VOI) of the rollouts and choosing a rollout that
maximizes the VOI. VOI of a rollout can be computed when the sample
distribution of an action is known up to the parameters, such as the
normal distribution with an unknown mean and/or
variance. Alternatively, the value of information can be estimated
from the set of samples, and the need to assume a particular shape of
the distribution can be lifted. In one realization of the latter
approach the VOI of performing an action is estimated as an
\emph{upper bound on the value of perfect information about an arm,
divided by the number of pulls of the arm}.
\begin{eqnarray}
VOI_\alpha&\approx&\frac {\overline X_\beta} {n_\alpha+1}
\exp\left(-2(\overline X_\alpha - \overline X_\beta)^2 n_\alpha\right)\\
VOI_i&\approx&\frac {1-\overline X_\alpha} {n_i+1}
\exp\left(-2(\overline X_\alpha - \overline X_i)^2 n_i\right),\; i\ne\alpha\nonumber\\
\mbox{where }\alpha&=&\arg\max_i \overline X_i\nonumber\\
             \beta&=&\arg\max_{i,\,i\ne\alpha} \overline X_i\nonumber
\end{eqnarray}
Early experiments with this approach demonstrated
significantly lower simple regret on a wide range of problems.

@@

\begin{figure}[h!]
  \begin{minipage}[c]{1.0\linewidth}
    \centering
    \includegraphics[scale=1.0]{onelevel-tree.pdf}\\
    a. search tree
    \vspace{1em}
  \end{minipage}
  \begin{minipage}[c]{1.0\linewidth}
    \centering
    \includegraphics[scale=0.5]{flat-trilevel-k=64-uqb=8.pdf}\\
    b. regret vs. number of samples
  \end{minipage}
  \caption{Simple regret in MAB}
  \label{fig:mab-simple-regret}
\end{figure}

@@

\begin{figure}[h!]
  \begin{minipage}[b]{1.0\linewidth}
    \centering
    \includegraphics[scale=0.45]{costs-size=6-group=median.pdf}\\
    a. median cost
    \vspace{1em}
  \end{minipage}
  \begin{minipage}[b]{1.0\linewidth}
    \centering
    \includegraphics[scale=0.45]{costs-size=6-group=minimum.pdf}\\
    b. minimum cost
  \end{minipage}
  \caption{The sailing domain, $6\times 6$ lake, cost vs. number of rollouts}
  \label{fig:sailing-cost-vs-nsamples}
\end{figure}

@@

\begin{figure}[h!]
  \centering
  \includegraphics[scale=0.45]{costs-size=6-nsamples=397.pdf}\\
  a. 397 rollouts\\
  \vspace{1em}
  \includegraphics[scale=0.45]{costs-size=6-nsamples=1585.pdf}\\
  c. 1585 rollouts
  \caption{The sailing domain, $6\times 6$ lake, cost vs. factor}
  \label{fig:sailing-cost-vs-factor}
\end{figure}

Figure~\ref{fig:sailing-lake-size} shows the cost vs. the exploration
factor for lakes of different sizes. The relative difference between
the sampling schemes becomes more prominent when the lake size
increases.
\begin{figure}[h!]
   \centering
   \includegraphics[scale=0.45]{costs-size=6-nsamples=397.pdf}\\
   b. $6\times 6$ lake \\
   \vspace{1em}
   \includegraphics[scale=0.45]{costs-size=10-nsamples=397.pdf}\\
   c. $10\times 10$ lake
  \caption{The sailing domain, 397 samples, cost vs. factor}
  \label{fig:sailing-lake-size}
\end{figure}

Figure~\ref{fig:sailing-rcq-vs-factor} compares QCT with UCT with a
different exploration factor at the root (CCT). $\alpha$ for the rest of
the steps was chosen to ensure the best performance from earlier
experiments on the domain. Both algorithms exhibit similar dependency,
but as the number of samples grows, QCT achieves smaller average
regrets and is less sensitive to the choice of the value for $\alpha$ at
the first step.
\begin{figure}[h!]
  \centering
  \includegraphics[scale=0.45]{rcq-size=6-nsamples=397.pdf}\\
  b. 397 rollouts\\
  \vspace{1em}
  \includegraphics[scale=0.45]{rcq-size=6-nsamples=1585.pdf}\\
  d. 1585 rollouts\\
  \caption{The sailing domain, log vs. sqrt, $6\times 6$ lake}
  \label{fig:sailing-rcq-vs-factor}
\end{figure}

@@

The MTCS SR+CR scheme presented in the paper differs
from UCT at the first step of the rollout, when the `simple' selection
regret is minimized instead of the cumulative regret. Both the
theoretical analysis and the empirical evaluation provide evidence for
better general performance of the proposed scheme.
